---
title: "437Project"
author: "ChenxinranShen (1004380092)"
date: "23/03/2020"
output:
  pdf_document: default
  html_document: default
---
# Appendix
Chenxinran Shen (1004380092)

## Preparation
```{r new}
library(dplyr)
library(MVN)
```
```{r start}
# import data
country<-read.csv("happiness2017.csv")
```
## Q2. 
First remove all rows with missing value, then random 100 sample from country without replacement.
```{r Q2 random sample of 100 countries}
# 1.random 100 sample without replacement,no missing value
set.seed(0092)
country<-na.omit(country)
head(country)

country_sample <- sample_n (country, 100,replace = FALSE)
#summary(country_sample)
```
Second, check the nomality of sample using functions from MVN package, by drawing chi-square plot. under the method of Henze-Zirkler’s test. 
Third, remove the outliers from the sample. Transform variables using square root transformation. Then remove the outliers in transformed data set.
Check nomality using chi-square plot again, we can see the transformed data is distributed as MVN right now.
```{r Q2}
# 2 nomality + outlier, remove outlier
country_sample_2<-country_sample[,-1]
par(mfrow=c(2,2))
result_oritinal <- mvn(data = country_sample_2, mvnTest = "hz", multivariateOutlierMethod = "quan")

# 3.1. remove outliers
result_noOutlier <- mvn(data = country_sample_2, mvnTest = "hz", multivariateOutlierMethod = "quan",showNewData = TRUE)
#result_noOutlier$newData+1
# 3.2. transform data
result_tran<- sqrt(result_noOutlier$newData)
# 3.3. check transformed data with no outlier's nomality
result_tran_noOutlier <- mvn(data = result_tran, mvnTest = "hz", multivariateOutlierMethod = "quan",showNewData = TRUE)

#result_tran_noOutlier
result <-mvn(result_tran_noOutlier$newData,multivariateOutlierMethod = "quan")[1]
new_sample <- result_tran_noOutlier$newData # tansformed data with no outliers
result

```
## Q3
Fit a linear regression model on original sampled data.
```{r Question 3 LM}
set.seed(0092)
# 1. Fit a linear model for the response variable- happiness score using all the original explanatory variables.
lmod<-lm(Ladder~Social+HLE+Freedom+Generosity+Corruption+Positive+Negative+gini,data = new_sample)
summary(lmod)


# only social and HLE have a small p value

# 2. Check whether the assumptions of this model are satiisfieded. 
par(mfrow=c(2,2))
plot(lmod, main = "                                            0092")
abline(lmod)

# 3. Report the `Adjusted R-squared' value from the R summary(lm(:::)) output
```

## Q4
```{r Question 4 PC/1}
set.seed(0092)
# 1. since the data set is normal ottain standardized variables
X= as.matrix(new_sample)  
x.bar = apply(X,2,mean)
S= cov(X)

Z=X
for(i in 1:ncol(Z)){
  Z[,i] = (X[,i]-x.bar[i])/sqrt(diag(S)[i])
}

# 2. From correlation matrix, get eigen value and eigen vectors
R = cor(X)
cov(Z)  # they should be the same!

# obtain eigenvalues and eigenvectors of R, from correlation matrix
Val.new = eigen(R)$values
#round(Val.new ,2)

Vec.new = eigen(R)$vectors
rownames(Vec.new) = colnames(X)
colnames(Vec.new) = c("PC1", "PC2", "PC3", "PC4", "PC5", "PC6","PC7","PC8","PC9","PC10")
#round(Vec.new ,2)

# 3. Get Principle Component values

W.new = X  # just to create a data matrix of the same size of X
colnames(W.new) = c("PC1", "PC2", "PC3", "PC4", "PC5", "PC6","PC7","PC8","PC9","PC10")

# now fill in the entries by calculating sample PCs

for(i in 1:ncol(new_sample)){ # PC's
 	for(j in 1:nrow(new_sample)){
	W.new[j,i] = Vec.new[,i] %*% Z[j,]   # no need to center when using normalized PCCs 
 }}

# Show Principal Components have zero correlation:

plot(data.frame(W.new),main = "Check Correlation between Principal Components 0092") 

# 4. How many components should we keep? 
# screeplot:

#plot(Val.new, type="b", pch=99, xlab="",ylab="Variances")  # suggests keeping the first 4 or 5 PCs.

# Proportion of variation explained by each PC:

round( Val.new/sum(Val.new),3)  

# If you like, you can use built-in functions in R for a summary:

summary(prcomp(Z))
screeplot(prcomp(Z),npcs = 10, type = "lines",main = "Screeplot for Happiness Data 0092") # choose 4
```
```{r Q4/1.2}
# 5. Regression with all standardized PCs as the explanatory variables

PC.model.new.1 <- lm(new_sample$Ladder ~  W.new[,1] + W.new[,2] + W.new[,3] + W.new[,4] + W.new[,5]+ W.new[,6]+W.new[,7]+W.new[,8]+W.new[,9])
summary(PC.model.new.1) 

# since principle complenent 1,4,8,9 have smallest p-value. use them to buile a new linear regression
PC.model.new.2 <- lm(new_sample$Ladder ~  W.new[,1] + W.new[,4]  + W.new[,9]+ +W.new[,8])
summary(PC.model.new.2)
PC.model.new.1$df.residual

```


